---
title: "ESM 204 - HW 4"
author: "Janelle Chistensen, Michael Wells & Rachel Rhodes"
date: "5/25/2020"
output: html_document
---

# Climate Change Model

The following climate-economic model assesses......


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)
```

```{r}
#-------------------------------
library(tidyverse)
library(janitor)
library(stargazer)
library(knitr)
library(kableExtra)
library(data.table)
library(purrr)
library(wesanderson)
##-----------------------
```

#### Model Parameters
```{r, include=TRUE,echo=TRUE}
#------------------------
#Variables
delta <- as.numeric(0.005)
eta <- as.numeric(0.5)
g <- as.numeric(0.01)
beta <- as.numeric(0.05)
y <- as.numeric(0:200)

##-----------------------
#Equation 1: Temperature over time
temp <- function(t,T){
  tau <- min(T*t/100,T)
  return(tau)
}

##-----------------------
#Equation 2: Activity retained over time 
K <- function(temp){
  output <- exp(-beta*(temp^2))
  return(output)
}

##-----------------------
#Equation 3: Total consumption over time
C <- function(K,t){
  output <- K*exp(g*t)
  return(output)
}

#-------------------------------
#Equation 4: Society's utility 
U <- function(C){
  output <- (C^(1-eta)/(1-eta))
  return(output)
}
```


```{r}
#-----------------------------------------
#Create Dataframe

df <- data.frame("year"= y) %>%
  group_by(year) %>% 
  mutate(temp_0 = temp(year,0),
         K_0=K(temp(year,0)),
         C_0=C(K(temp(year,0)),year),
         U_0=U(C(K(temp(year,0)),year)),
         temp_2 = temp(year,2),
         K_2=K(temp(year,2)),
         C_2=C(K(temp(year,2)),year),
         U_2=U(C(K(temp(year,2)),year)),
         temp_8 = temp(year,8),
         K_8=K(temp(year,8)),
         C_8=C(K(temp(year,8)),year),
         U_8=U(C(K(temp(year,8)),year)))


```


#### PLOTS
```{r}
#Plot temps over time

df_temp <- df %>% 
  select(year,temp_0, temp_2, temp_8) %>% 
  pivot_longer(-year, names_to = "Scenario", values_to = "temperature") %>% 
  mutate(Scenario = replace(Scenario, Scenario == "temp_0", "No Climate Change")) %>%
  mutate(Scenario = replace(Scenario, Scenario == "temp_2", "Modest Climate Change")) %>%
  mutate(Scenario = replace(Scenario, Scenario == "temp_8", "Extreme Climate Change")) 


ggplot(df_temp, aes(x = year, y = temperature, color = Scenario))+
  geom_line(aes(x = year,
                 y = temperature,
                 color = Scenario))+
  scale_color_manual(values = wes_palette("Darjeeling1"))+
  labs(x="Years",y="Temperature")+
  ggtitle("Temperature Over 200 Years")+
  theme_minimal()
```
***Figure 1*** *Predicted temperature over 200 years under three different climate change scenarios: no climate change (T=0), modest climate change (T=2), and extreme climate change (T=8).*



```{r}
#------------------------------------------
#Plot consumption 

df_consump <- df %>% 
  select(year,C_0, C_2, C_8) %>% 
  pivot_longer(-year, names_to = "Scenario", values_to = "Consumption") %>% 
  mutate(Scenario = replace(Scenario, Scenario == "C_0", "No Climate Change")) %>%
  mutate(Scenario = replace(Scenario, Scenario == "C_2", "Modest Climate Change")) %>%
  mutate(Scenario = replace(Scenario, Scenario == "C_8", "Extreme Climate Change")) 

ggplot(df_temp, aes(x = year, y = Consumption, color = Scenario))+
  geom_line(aes(x = year,
                 y = temperature,
                 color = Scenario))+
  scale_color_manual(values = wes_palette("Darjeeling1"))+
  labs(x="Years",y="Total Consumption")+
  ggtitle("Total Consumption Over 200 Years")+
  theme_minimal()
```


***Figure 2*** *Predicted total consumption over 200 years under three different climate change scenarios: no climate change (T=0), modest climate change (T=2), and extreme climate change (T=8).*



```{r}
#------------------------------------------
#Plot undiscounted utility

df_utility<- df %>% 
  select(year,U_0, U_2, U_8) %>% 
  pivot_longer(-year, names_to = "Scenario", values_to = "Utility") %>% 
  mutate(Scenario = replace(Scenario, Scenario == "U_0", "No Climate Change")) %>%
  mutate(Scenario = replace(Scenario, Scenario == "U_2", "Modest Climate Change")) %>%
  mutate(Scenario = replace(Scenario, Scenario == "U_8", "Extreme Climate Change")) 

ggplot(df_temp, aes(x = year, y = Utility, color = Scenario))+
  geom_line(aes(x = year,
                 y = temperature,
                 color = Scenario))+
  scale_color_manual(values = wes_palette("Darjeeling1"))+
  labs(x="Years",y="Utility")+
  ggtitle("Society's Utility Over 200 Years")+
  theme_minimal()

```


***Figure 3*** *Predicted undiscounted utility to society over 200 years under three different climate change scenarios: no climate change (T=0), modest climate change (T=2), and extreme climate change (T=8).*

```{r, include=TRUE, results="hide"}
#-------------------------------------------

#2a.Present value over the next 200 years with moderate climate change (t=4.4) compared to no climate change (T=0)

r <- delta +eta*g

df_4.4 <- data.frame("year"= y) %>%
  group_by(year) %>% 
  mutate(temp_4.4 = temp(year,4.4),
         K_4.4=K(temp(year,4.4)),
         C_4.4=C(K(temp(year,4.4)),year),
         U_4.4=U(C(K(temp(year,4.4)),year)),
         pv_u_4.4=U_4.4/((1+r)^year),
         temp_0 = temp(year,0),
         K_0=K(temp(year,0)),
         C_0=C(K(temp(year,0)),year),
         U_0=U(C(K(temp(year,0)),year)),
         pv_u_0=U_0/((1+r)^year))

npv_pv_u_4.4 <- sum(df_4.4$pv_u_4.4)
npv_pv_u_0 <- sum(df_4.4$pv_u_0)

L <- ((npv_pv_u_4.4/npv_pv_u_0)-1)*100

# Create a table
table_4.4 <- data.frame(scenario =c("No Climate Change", "Moderate Climate Change", "Percent Loss"),
                        net_present_value = c(npv_pv_u_0, npv_pv_u_4.4, L))


```

#### ANALYSIS 

##### 2.A Net Present Value & Percent Loss 
As shown in Table 1, the present value utility over 200 years with moderate climate change (T = 4.4) is `r npv_pv_u_4.4` compared to the present value utility with no climate change (T = 0) which is `r npv_pv_u_0`. There is 

[insert Table]
```{r, include=TRUE, results="hide"}
#--------------------------------

#2b: new variables w/ 10% increase


T_10pct <- 4.4*1.1
g_10pct <- g*1.1
eta_10pct <- eta*1.1
beta_10pct <- beta*1.1

#-----------------------------------------
#2b cont'd: 10% change in T
df_T_10pct <- data.frame("year"= y) %>%
  group_by(year) %>% 
  mutate(temp_T_10pct = temp(year,T_10pct),
         K_T_10pct=K(temp(year,T_10pct)),
         C_T_10pct=C(K(temp(year,T_10pct)),year),
         U_T_10pct=U(C(K(temp(year,T_10pct)),year)),
         pv_u_T_10pct=U_T_10pct/((1+r)^year),
         temp_0 = temp(year,0),
         K_0=K(temp(year,0)),
         C_0=C(K(temp(year,0)),year),
         U_0=U(C(K(temp(year,0)),year)),
         pv_u_0=U_0/((1+r)^year))

npv_pv_u_T_10pct <- sum(df_T_10pct$pv_u_T_10pct)
npv_pv_u_0 <- sum(df_T_10pct$pv_u_0)

L_T10pct <- ((npv_pv_u_T_10pct/npv_pv_u_0)-1)*100

#-------------------------------------------

#2b cont'd: 10% change in g

r_g_10pct <- delta +eta*g_10pct

C_g10pct <- function(K,t){
  output <- K*exp(g_10pct*t)
  print(output)
}

df_g_10pct <- data.frame("year"= y) %>%
  group_by(year) %>% 
  mutate(temp_4.4 = temp(year,4.4),
         K_4.4=K(temp(year,4.4)),
         C_4.4=C_g10pct(K(temp(year,4.4)),year),
         U_4.4=U(C_g10pct(K(temp(year,4.4)),year)),
         pv_u_4.4=U_4.4/((1+r_g_10pct)^year),
         temp_0 = temp(year,0),
         K_0=K(temp(year,0)),
         C_0=C_g10pct(K(temp(year,0)),year),
         U_0=U(C_g10pct(K(temp(year,0)),year)),
         pv_u_0=U_0/((1+r_g_10pct)^year))

npv_pv_u_4.4_g10pct <- sum(df_g_10pct$pv_u_4.4)
npv_pv_u_0_g10pct <- sum(df_g_10pct$pv_u_0)

L_g10pct <- ((npv_pv_u_4.4_g10pct/npv_pv_u_0_g10pct)-1)*100

#-----------------------------------------
#2b cont'd: 10% change in eta

r_eta10pct <- delta +eta_10pct*g

U_eta10pct <- function(C){
  output <- (C^(1-eta_10pct)/(1-eta_10pct))
  print(output)
}

df_eta10pct <- data.frame("year"= y) %>%
  group_by(year) %>% 
  mutate(temp_4.4 = temp(year,4.4),
         K_4.4=K(temp(year,4.4)),
         C_4.4=C(K(temp(year,4.4)),year),
         U_4.4=U_eta10pct(C(K(temp(year,4.4)),year)),
         pv_u_4.4=U_4.4/((1+r_eta10pct)^year),
         temp_0 = temp(year,0),
         K_0=K(temp(year,0)),
         C_0=C(K(temp(year,0)),year),
         U_0=U_eta10pct(C(K(temp(year,0)),year)),
         pv_u_0=U_0/((1+r_eta10pct)^year))

npv_pv_u_4.4_eta10pct <- sum(df_eta10pct$pv_u_4.4)
npv_pv_u_0_eta10pct <- sum(df_eta10pct$pv_u_0)

L_eta10pct <- ((npv_pv_u_4.4_eta10pct/npv_pv_u_0_eta10pct)-1)*100

#-----------------------
#2b cont'd:  10% change in beta

K_beta10pct <- function(temp){
  output <- exp(-beta_10pct*(temp^2))
  print(output)
}

df_beta10pct <- data.frame("year"= y) %>%
  group_by(year) %>% 
  mutate(temp_4.4 = temp(year,4.4),
         K_4.4=K_beta10pct(temp(year,4.4)),
         C_4.4=C(K_beta10pct(temp(year,4.4)),year),
         U_4.4=U(C(K_beta10pct(temp(year,4.4)),year)),
         pv_u_4.4=U_4.4/((1+r)^year),
         temp_0 = temp(year,0),
         K_0=K(temp(year,0)),
         C_0=C(K(temp(year,0)),year),
         U_0=U(C(K(temp(year,0)),year)),
         pv_u_0=U_0/((1+r)^year))

npv_pv_u_4.4_beta10pct <- sum(df_beta10pct$pv_u_4.4)
npv_pv_u_0_beta10pct <- sum(df_beta10pct$pv_u_0)

L_beta10pct <- ((npv_pv_u_4.4_beta10pct/npv_pv_u_0_beta10pct)-1)*100

show(L_beta10pct)

# Percent change in L
L_eta <- ((L_eta10pct/L)-1)*100
L_beta <- ((L_beta10pct/L)-1)*100
L_g <- ((L_g10pct/L)-1)*100
L_T <- ((L_T10pct/L)-1)*100
```

```{r}
#-------------------------------
#2c 

#theta^(1-eta) * npv_pv_u_4.4 = npv_pv_u_0

theta <- function(npv_pv_u_4.4, npv_pv_u_0)
{
  output <- 1 - (npv_pv_u_4.4/npv_pv_u_0)^((1-eta)^-1)
  return(output)
}

theta(npv_pv_u_4.4,npv_pv_u_0)

```


```{r}
#--------------------------------
#2d. Uncertainty
## T = 2 w probability of 0.2
## T = 4 w probability of 0.5
## T = 6 with a probability of 0.3

## p * utility(T=2) + p * utillity(T=4) + p * utility(T=6)

df_uncertain <- data.frame("year"= y) %>%
  group_by(year) %>% 
  mutate(temp_2 = temp(year,2),
         K_2=K(temp(year,2)),
         C_2=C(K(temp(year,2)),year),
         U_2=U(C(K(temp(year,2)),year)),
         pv_u_2=U_2/((1+r)^year),
         temp_4 = temp(year,4),
         K_4=K(temp(year,4)),
         C_4=C(K(temp(year,4)),year),
         U_4=U(C(K(temp(year,4)),year)),
         pv_u_4=U_4/((1+r)^year),
         temp_6 = temp(year,6),
         K_6=K(temp(year,6)),
         C_6=C(K(temp(year,6)),year),
         U_6=U(C(K(temp(year,6)),year)),
         pv_u_6=U_6/((1+r)^year))

npv_pv_u_2 <- sum(df_uncertain$pv_u_2)
npv_pv_u_4 <- sum(df_uncertain$pv_u_4)
npv_pv_u_6 <- sum(df_uncertain$pv_u_6)

npv_pv_u_uncertain <- (0.2 * npv_pv_u_2) + (0.5 * npv_pv_u_4) + (0.3 * npv_pv_u_6)

theta(npv_pv_u_uncertain, npv_pv_u_0)



```

